---
title: "1-21-21_ClassNotes"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 


<h1>An Introduction to `R`</h3>
```{r}
seq(1,10)

seq(6,15)

seq(-10,-1)

seq(5,-4)

seq(1.5, 10.5)
```
the sequence can print out the numbers up to 10 if it is incremented by 1 from 1.5 to 10 ie:
```{r}
seq(1.5, 10)
```
<h3>help</h3>
```{r}
help(seq)
```
can specify the value of each by typing the name or you can type it in the same order
```{r}
seq(from=1, to =1, by ((to -from) / (lenth.out -1)), length.out = NULL, along.with = NULL)

seq(1,10,2)
# is the same as
seq(from=1, to=10, by=2)
```
<h2>Vector `c()`</h2>
`c()` will concatenate a list of values into a vector
```{r}
x = c(1,2,3,4,5,6)
x

x = c(1,4,3)
x = c("blue", "red", "yellow")

```
indexing a vector starts at 1 in `R`
```{r}
# not zero based indexing
x = c(1,2,3,4,5,6,7,8,9,10)
x1 = x[1]
x2 = x[3:5]
x1
x2

```
getting the length, min or max of a vector
```{r}
length(x)
min(x)
max(x)
```
calculating central tendency of values in a vector, ie. mean and median
```{r}
mean(x)
median(x)
```
calculating spread of values of a vector using the standard deviating or the variance 
```{r}
#standard deviating
sd(x)

# variance
var(x)
```
<h2>Creating a *Matrix*</h2> 
the function `matrix()` creates a two-dimensional data structure

the following would set `A` as a two by two matrix
```{r}
A = matrix(data=c(1,2,3,4), nrow=2, ncol=2)
A
```
<h3> indexing data in a matrix</h3>
suppose we have the following matrix:
```{r}
A = matrix(seq(1,16), 4,4)
A
```
we can select the element in the second row and the fourth column:
```{r}
A[2,4] # A[row,column]
```
you can also select multiple rows and columns in a matrix. it will create a new matrix as it's output
```{r}
A
b = A[c(1,3), c(2,4)] # A[row, column]
b
```
can get all rows or columns by leaving the space blank
```{r}
A[1:2, ]
```
*remove* a row or column with a negative sign
```{r}
A
A[-c(1,3),]
```
`dim()` will return he matrix's dimensions
```{r}
dim(A)
```
also: min, max, median, and sd are also available
```{r}
min(A)
max(A)
median(A)
sd(A)
```
<h3>Manipulating all values of a Vector or Matrix</h3>
suppose we have a vector: `x = c(1,2,3,4,5,6,7,8,9,10)`. We can manipulate all of its values at once. ex:
```{r}
x = c(1,2,3,4,5,6,7,8,9,10)

# Divide all of its values by two
x/2
```
vector arithmetic
```{r}
x = c(1,2,3,4,5,6,7,8,9,10)
y = c(10,9,8,7,6,5,4,3,2,1)

x+y
```
can also be done with matrices, but must be careful to make sure they are the same dimensions or else it will loop around and continue the arithmetic

<h2>Installing Packages</h2>
will be using the CRAN packages, will be working with `tidyverse` package in this lecture
```{r}
install.packages("tidyverse")
```


