---
title: "Cross Validation"
output: html_notebook
---

# Cross Validation

## Syntac for `cv.glm()`

- the function `cv.glm()` calculates estimated k-fold cross-validation prediction error for generalized liner models:
- `cv.glm(data, glmfit, cost, k)`
- `data`: matrix or data from containing the full dataset
- `glmfit`: object containing results of `glm()` fitted to `data`
- `cost`: function specifying cost function (default is MSE)
- `K`: number of folds to split data into (default is n)
- thus, we don't need to set `cost` at all, or `K` if we want to use LOOCV (leave-one-out Cross validation).

#### still using `Advertising` dataset for linear regression
```{r}
library(ISLR)
library(boot)
library(tidyverse)
attach(Advertising)
```

## Perfroming LOOCV with linear regression
- we just used `glm()` to fit linear regression model to predict `sales` from `TV`, `radio`, and `newspaper`:
```{r}
glm.fit = glm(sales ~ TV + radio + newspaper)

# to perform LOOCV, we just type:
cv.err = cv.glm(Advertising, glm.fit)

# then to view cross-validation test error and training error:
cv.err$delta

# standard deviation
sd(sales)
```

- perform k-fold C with k=5
- when K=5, we have 100 observations, we split it up into 5 folds with 20 observations in each. Then we go through those folds one at time and each time we are using a different fold for testing. It randomly splits the data in one of the five folds. 
```{r}
set.seed(1)
cv.err = cv.glm(Advertising, glm.fit, K=5)

cv.err$delta
```
- LOOCV (leave-one-out crass validation) is very resource intensive and that is why it's not used very much
```{r}
set.seed(2)
cv.err = cv.glm(Advertising, glm.fit, K=10)

cv.err$delta
```

## Review: Performance of three classifiers
- Because we used the same training and test sets for all three classifiers, we were able to directly compare their performance:
- KNN (𝐾 = 10*) test accuracy rate = 0.967, test error rate = 0.033 
- Logistic regression test accuracy rate = 0.967, test error rate = 0.033 
- LDA test accuracy rate = 0.970, test error rate = 0.03
- Thus, in this specific scenario, LDA appeared to be the most accurate. 
-- But with CV, we get a lower test error rate for logistic regression!

## Summary of concepts covered in Lecture 9
- Evaluating performance of statistical learning methods is critical. 
-- MSE for regression and proportion of errors for classification
- Training error (and fit) are important to assess, but not good enough. 
-- Need to evaluate test error!
- When there is no test data, we use cross-validation: LOOCV or 𝑘-fold CV. 
-- Gives us mean MSE or mean proportion of errors
- LOOCV is superior to 𝑘-fold CV, BUT can be computationally expensive! 
-- Typically use 𝑘-fold CV with 𝑘 = 5 or 𝑘 = 10.