---
title: "Homework 3"
output:
  html_document:
    df_print: paged
---

## will be using the College dataset in the ISLR package
- loading the package and dataset in 
```{r}
library(ISLR)
summary(College)
library(ggplot2)
library(tidyverse)
# install.packages("class")
# set.seed(1) 
#library(class)
#attach(College)
#attach(Default)
```

## make predictions about a college’s out-of-state tuition from all other variables in College

```{r}
lm.fit = lm(formula = Outstate ~ Private + Apps + Accept + Enroll + Top10perc + Top25perc + F.Undergrad + P.Undergrad + Room.Board + Books + Personal + PhD + Terminal + S.F.Ratio + perc.alumni + Expend + Grad.Rate, data = College)
#lm.fit

summary(lm.fit)
```

```{r}
ggplot(data = lm.fit, mapping = aes(x=(Apps + Accept + Enroll + Top10perc + Top25perc + F.Undergrad + P.Undergrad + Room.Board + Books + Personal + PhD + Terminal + S.F.Ratio + perc.alumni + Expend + Grad.Rate), y=Outstate)) + geom_jitter() + geom_smooth(method='lm', formula= y~x) + theme_bw()
```

## 2
- need to break observations that are tied as nearest neighbors in a reproducible way. 
```{r}
library(class)
```
 

#### a

- slide 81 of classification lecture. 
- about 34 min of march 17th lecture she goes over it. 
```{r}
attach(College)
train = seq(1,500)
test = seq(501, 777)
train.x = cbind(College[train, 2:18])
test.x = cbind(College[test, 2:18])
train.y = Private[train]
head(test.x)
```

- need to create a vector class of labels for these features.
```{r}
#train.y <- default[train]
#train.y
```

#### b

```{r}
# k=1
set.seed(1) 
knn.pred.1 = knn(train.x, test.x, train.y, k=1)
summary(knn.pred)
# k=10
set.seed(1) 
knn.pred.10 = knn(train.x, test.x, train.y, k=10)
summary(knn.pred.10)
# k=100
set.seed(1) 
knn.pred.100 = knn(train.x, test.x, train.y, k=100)
summary(knn.pred.100)
```

#### c
- The goal is to compare these predicted classes to their true classes.
- To make comparisons easier, let’s create a vector of true test classes
- table, predication in rows and observations in the columns
```{r}
test.y = Private[test]
table(knn.pred.1, test.y)
table(knn.pred.10, test.y)
table(knn.pred.100, test.y)
```

#### d
 - using the mean function to provide the test accuracy
- *table size is still off, gave me an answer though.... 
```{r}
mean(knn.pred.1 == test.y)
mean(knn.pred.10 == test.y)
mean(knn.pred.100 == test.y)
```
 
#### e
- Provide test error rates for your KNN classifier when K=1, 10, and 100
```{r}
1 - mean(knn.pred.1 == Private[test])
1 - mean(knn.pred.10 == Private[test])
1 - mean(knn.pred.100 == Private[test])
```

