---
title: "Homework 1"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

### Robert Haire
```{r}
install.packages("ISLR")
```
<ul>
  <li>1.a) How many colleges are in the dataset?</li>
    <ul>
      There are 777 different universities in this dataset
    </ul>
  <li>1.b) How many features are there for each college?
    <ul>
      There are 18 different features for each college.
    </ul>
  </li>
  <li>1.c) Which features is(are) categorical?
    <ul>
      
    </ul>
  </li>
  <li>1.d) Which feature(s) is(are) numerical?
    <ul>
    #answer
    </ul>
  </li>
  <li>1.e) How many colleges are private?
    <ul>
    #answer
    </ul>
  </li>
  <li>1.f) What is the mean graduation rate across colleges?
    <ul>
    #answer
    </ul>
  </li>
  <li>1.g) What is the maximum number of undergraduate students at a college? <br>Hint: Make sure that you include both full-time and part-time students in your calculation. 
    <ul>
    #answer
    </ul>
  </li>
</ul>

```{r}
read.csv(college)
```


