---
title: "Homework 1"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

### Robert Haire
```{r}
install.packages("ISLR")
```


```{r}
library(ISLR)

```

```{r}
help(College)
```

```{r}
College
```

```{r}
summary(College)
```

```{r}
maxF_UnderGrad = max(College$F.Undergrad)
maxP_UnderGrad = max(College$P.Undergrad)
maxUnderGrad = maxF_UnderGrad + maxP_UnderGrad
maxUnderGrad
```
# 2
Second, we will examine the relationship between the percentage of students from the top 10% of their high school class and the out-of-state tuition at a college. 
## 2.a 
Create a scatterplot with the percentage of students from the top 10% of their high school class on the x axis and the out-of-state tuition on the y axis. Overlay the points with a smoothed line and 95% confidence bands. Remember to avoid overplotting. 

```{r}
ggplot(data = College) + geom_point(mapping = aes(x = Top10perc, y = Outstate)) + geom_smooth(mapping = aes(x = Top10perc, y = Outstate), level = 0.95)
```

# 3
## 3.a
[10 points] Using different colored points for public and private colleges, create a scatterplot with the percentage of students from the top 10% of their high school class on the x axis and the out-of-state tuition on the y axis. Overlay each set of points with a smoothed line of the same color with 95% confidence bands. Remember to avoid overplotting. 
```{r}
ggplot(data = College, mapping = aes(x = Top10perc, y = Outstate)) + geom_point(mapping = aes(color = class)) + geom_smooth()
```

```{r}
ggplot(data = College, mapping = aes(x = Top10perc, y = Outstate, color = class)) + geom_point() + geom_smooth()
```



