---
title: "2.3 lab intro notes"
output: html_notebook
---

To run a function called `funcname`, we type `funcname(input1, input2)`, where the inputs (or arguments) `input1` and `input2` tell `R` how to run the function. 

```{r}
x <- c(1,2,3,5)
x

```
dont need to use <- can use equal sign instead
```{r}
x = c(1,6,2)
x
y= c(1,4,3)
y
```
add two sets of numbers together
x = (x1,x2,x2, ...,xn)
y = (y1,y2,y3, ..., yn)

x + y = (x1+y1, x2+y2, x3+y3, ..., xn+yn)
```{r}
length(x) # displays length of x
length(y) # displays length of y
x+y # adds each together via their position
```
The `ls()` functions allows us to look at a *list* of all the objects, such as data functions, that we have saved so far. 
The `rm()` function can be used to *delete* any that we don't want. 
```{r}
ls()
```

```{r}
rm(x,y)
ls()
```
interesting to see that `z` is still there, never did delete it. still kept in memory, lets try to remove it...
```{r}
rm(z)
ls()
```
not it says `0`, need to keep in mind when doing large sets of data to remove any unneeded variable to not slow down the compile

can also *remove all* objects at once:
```{r}
rm(list=ls())
```

The `matrix()` function can be used to create a matrix of numbers
```{r}
?matrix() # calls help function
```

*Usage*
matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE,
       dimnames = NULL)

as.matrix(x, ...)
## S3 method for class 'data.frame'
as.matrix(x, rownames.force = NA, ...)

is.matrix(x)

*Arguments*
data	
an optional data vector (including a list or expression vector). Non-atomic classed R objects are coerced by as.vector and all attributes discarded.

nrow	
the desired number of rows.

ncol	
the desired number of columns.

byrow	
logical. If FALSE (the default) the matrix is filled by columns, otherwise the matrix is filled by rows.

dimnames	
A dimnames attribute for the matrix: NULL or a list of length 2 giving the row and column names respectively. An empty list is treated as NULL, and a list of length one as row names. The list can be named, and the list names will be used as names for the dimensions.

x	
an R object.

...	
additional arguments to be passed to or from methods.

rownames.force	
logical indicating if the resulting matrix should have character (rather than NULL) rownames. The default, NA, uses NULL rownames if the data frame has ‘automatic’ row.names or for a zero-row data frame.

```{r}
x = matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
```
We can omit typing `data=`, `nrow=`, and `ncol=` in the `matrix()` and just type
```{r}
x=matrix(c(1,2,3,4),2,2)
x
```

By default, `R` populates matrices by column first but this can be adjusted by setting `byrow=TRUE` 
```{r}
matrix(c(1,2,3,4),2,2,byrow=TRUE)
```

`sqrt()` returns square root of each element of a vector or matrix.
`x^2` raises each element of `x` to the power of `2`
```{r}
sqrt(x)
x^2
```

The `rnorm()` function generates a *vector of random normal variables* with the first argument `n` the sample size. 
`cor()` calculates the relationship between each column and row
`sd=` standard deviation
```{r}
x=rnorm(50)
y=x+rnorm(50, mean=50, sd=.1)
cor(x,y)
```

`set.seed()`

rotissier chicken, salad american, chicken waffles? 